Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2013-12-21 paper.Rnw
WAval plot

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
\usepackage{titlesec, siunitx, graphicx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}   % keep floats in their place.
\usepackage{authblk}             % Listing Author affiliations
\usepackage{rotating}            % rotating figures
\usepackage[hypcap]{caption}

@@

\begin{document}
\maketitle

@@

In this paper we have describe a thorough validation of the \mb{} algorithm for
segmentation of the hippocampus and its subfields. First, we addressed the feasibility and
accuracy of whole hippocampus segmentation with a limited number of input
atlases \citep{MallarChakravarty2012} by performing a multi-fold experiment
using a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
dataset over a range of parameters.  We then performed a leave-one-out
validation to determine if hippocampal subfields can be accurately identified
using our \multiatlas framework.  To ensure that we have not over-fit our
parameters to the aging or neurodegenerative brain, we also applied our
segmentations to a dataset of individuals suffering from first episode
psychosis.  Finally, validated our algorithm using all of the data available in
the \adnidataset{} sample and compare our segmentations to the other
segmentations available through the ADNI informatics portal.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: volume plots
\begin{figure}[h!]
<<WAval-vol-plots,cache=T,fig.width=6,fig.height=6>>=
wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO images"="mb_bravo")

wa.xval.vols.lm = ddply(wa.xval.volumes, c("region", "measure"), function(df) {
   model = lm(volume ~ gold_0.3mm, df)
   pearson = cor(df$volume, df$gold_0.3mm, use='pairwise.complete.obs')
   data.frame(
     eqn  = lm_eqn(model), 
     corr = pearson)})

ggplot(wa.xval.volumes, aes(x=gold_0.3mm, y=volume, colour=measure)) + 
  facet_wrap(~region, ncol=3, scales='free') + 
  geom_point() + 
  geom_smooth(method='lm')  + 
  coord_equal() + 
  ylab(expression("Automated Unilateral Volume" (mm^3))) +
  xlab(expression("Manual Unilateral Volume" (mm^3))) +
  geom_text(aes(x=Inf-200, y=Inf-200, label=eqn), parse=TRUE, data=wa.xval.vols.lm, position="dodge") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
\begin{figure}
<<fig:ADNI1-SNT-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=5, fig.height=5>>=
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(subset(snt_xval_mean, (templates.mb*atlases) %% 2 == 1),
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',
               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.05)) + 
  scale_colour_hue(name="Number of Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text = element_text(size = 8))
@
  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
  SNT segmentations for 69 ADNI subjects vs. atlas and template library
  size, registration algorithm, and label fusion method. 
  Error bars indicate standard error.
  \label{fig:ADNI1-SNT-xval-k-mean}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
\usepackage{titlesec, siunitx, graphicx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}   % keep floats in their place.
\usepackage{authblk}             % Listing Author affiliations
\usepackage{rotating}            % rotating figures
\usepackage[hypcap]{caption}
\usepackage{lineno}

@@

\begin{document}
\maketitle
\linenumbers

@@

In this paper we have describe a thorough validation of the \mb{} algorithm for
segmentation of the hippocampus and its subfields. First, we address the feasibility and
accuracy of whole hippocampus segmentation with a limited number of input
atlases \citep{MallarChakravarty2012} by performing a multi-fold experiment
using a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
dataset over a range of parameters. We then perform a leave-one-out
validation to determine if hippocampal subfields can be accurately identified,
using the best parameters discovered in the first experiment.  To ensure that we
have not overfit our parameters to the aging or neurodegenerative brain, we also
apply \mb{} to a dataset of individuals suffering from first episode psychosis.
Finally, we validat our algorithm using all of the data available in the
\adnidataset{} sample and compare our segmentations to other popular
segmentation algorithms.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: volume plots
\begin{figure}[h!]
<<WAval-vol-plots,cache=T,fig.width=7,fig.height=6>>=
wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO images"="mb_bravo")

wa.xval.vols.lm = ddply(wa.xval.volumes, c("region", "measure"), function(df) {
   model = lm(volume ~ gold_0.3mm, df)
   pearson = cor(df$volume, df$gold_0.3mm, use='pairwise.complete.obs')
   data.frame(
     eqn  = lm_eqn(model), 
     corr = pearson)})
wa.xval.vols.lm$yoffset = c( 1225 - 0:2 * 40,
                             465  - 0:2 * 15, 
                             750  - 0:2 * 30,
                             325  - 0:2 * 18, 
                             1100 - 0:2 * 40 )

wa.xval.vols.lm$xoffset = c( rep(700, 3),
                             rep(250, 3),
                             rep(490, 3),
                             rep(140, 3), 
                             rep(570, 3) )

ggplot(wa.xval.volumes, aes(x=gold_0.3mm, y=volume, colour=measure)) + 
  facet_wrap(~region, ncol=3, scales='free') + 
  geom_point() + 
  coord_equal() + 
  geom_smooth(method='lm')  + 
  xlab(expression("Manual Unilateral Volume" (mm^3))) +
  ylab(expression("Automated Unilateral Volume" (mm^3))) +
  geom_text(aes(x=xoffset, y=yoffset, label=eqn), parse=TRUE,
    data=wa.xval.vols.lm, size=2,hjust=0) +
  theme(legend.direction = "horizontal", legend.position = "bottom")

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
\begin{figure}
<<fig:ADNI1-SNT-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=5, fig.height=5>>=
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(subset(snt_xval_mean, (templates.mb*atlases) %% 2 == 1),
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
  facet_grid(.~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',
               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.05)) + 
  scale_colour_hue(name="Number of Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text = element_text(size = 8))
@
  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
  SNT segmentations for 69 ADNI subjects vs. atlas and template library
  size, registration algorithm, and label fusion method. 
  Error bars indicate standard error.
  \label{fig:ADNI1-SNT-xval-k-mean}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
